---
title: 'Main Takeaways'
output:
  distill::distill_article:
    toc:false:
      theme: theme.css
favicon: "images/favicon.ico"
github-repo: childpovertyactionlab/DPD-incident-reports
---

```{r Set-Up Block, include = FALSE}
#### Libraries to load #####
library(tidyverse)
library(sf)
library(leaflet)
library(mapboxapi)
library(lubridate)
library(rio)
library(tidycensus)
library(htmltools)
library(arcgisbinding)
library(reactable)
library(CPALtools)

#datalib <- "C:/Users/micha/CPAL Dropbox/" # Michael Laptop
datalib <- "E:/CPAL Dropbox/" # Michael Desktop

#### Defining vgc and firearms #####
firearm <- c("Handgun", "Rifle","Firearm (Type Not Stated)", "Assault Weapon", "Unknown Type Gun", "Other/Unknown Gun", "Shotgun", "Other Gun", "Other Firearm")
vgc_type <- c("AGG ASSAULT - NFV", "MURDER & NONNEGLIGENT MANSLAUGHTER", "NEGLIGENT MANSLAUGHTER")

#### Importing Dallas Open Data portals and filtering into data frames #####
dpdIncidents <- import(paste0(datalib, "Data Library/City of Dallas/04_Public Safety/Dallas Police/Incidents/Incidents by Type/GroupA.csv")) %>%
  st_as_sf(coords = c(x = "x_coordinate", y = "y_coordinate"),
                                   crs = "ESRI:102738") %>%
  mutate(Date = as.Date(Date),
         Time = lubridate::hm(Time),
         Year = year(Date),
         Month = month(Date),
         Day = day(Date),
         division = str_to_title(division),
         crime_category = as.factor(nibrs_crime_category),
         crime = as.factor(nibrs_crime),
         Weekday = wday(Date),
         DateTime = format(paste(Date, Time), format="%Y-%m-%d %H:%M:%S"),
         hour = hour(Time),
         vgc_flag = ifelse(nibrs_crime %in% vgc_type & Firearm == 1, 1, 0),
         ToD = ifelse(hour >= 3 & hour < 7, "Early Morning",
                      ifelse(hour >= 7 & hour < 13, "Morning",
                             ifelse(hour >= 13 & hour < 18, "Afternoon",
                                    ifelse(hour >= 18 & hour < 23, "Evening",
                                           ifelse(hour >= 23 | hour < 3, "Night", "NA")
                                           )
                                    )
                             )
                      )
         ) %>%
  mutate(crime_title = ifelse(crime == "MURDER & NONNEGLIGENT MANSLAUGHTER", "Murder & Nonnegligent Manslaughter",
                              ifelse(crime == "VIOLENT GUN CRIME", "Violent Gun Crime",
                                     ifelse(crime == "AGG ASSAULT - NFV", "Aggravated Assault (NFV)",
                                            ifelse(crime == "ROBBERY-BUSINESS", "Robbery (Business)",
                                                   ifelse(crime == "ROBBERY-INDIVIDUAL", "Robbery (Individual)",
                                                          ifelse(crime == "BURGLARY-RESIDENCE", "Burglary (Residence)",
                                                                 ifelse(crime == "BURGLARY-BUSINESS", "Burglary (Business)",
                                                                        "ERROR"
                              )))))))) %>%
  st_transform(crs = 4269)

#unique(dpdIncidents$crime)

currentmonth <- as.Date(paste0(year(today()), "-", month(today()), "-01"))

last30 <- currentmonth-months(1)

vgc5year <- dpdIncidents %>%
  filter(vgc_flag == 1,
         Date >= as.Date("2016-01-01"))

vgc12months <- dpdIncidents %>%
  filter(vgc_flag == 1,
         Date >= as.Date("2021-01-01"))

vgc30days <- dpdIncidents %>%
  filter(vgc_flag == 1,
         Date >= last30 & Date < currentmonth)

AggAs <- dpdIncidents %>%
  filter(crime == "AGG ASSAULT - NFV",
         Date >= last30 & Date < currentmonth)

RobBus <- dpdIncidents %>%
  filter(crime == "ROBBERY-BUSINESS",
         Date >= last30 & Date < currentmonth)

RobInd <- dpdIncidents %>%
  filter(crime == "ROBBERY-INDIVIDUAL",
         Date >= last30 & Date < currentmonth)

BurRes <- dpdIncidents %>%
  filter(crime == "BURGLARY-RESIDENCE",
         Date >= last30 & Date < currentmonth)

BurBus <- dpdIncidents %>%
  filter(crime == "BURGLARY-BUSINESS",
         Date >= last30 & Date < currentmonth)

dpdMini <- dpdIncidents %>%
  filter(crime %in% c("AGG ASSAULT - NFV", "ROBBERY-BUSINESS", "ROBBERY-INDIVIDUAL", "BURGLARY-RESIDENCE", "BURGLARY-BUSINESS", "MURDER & NONNEGLIGENT MANSLAUGHTER")) %>%
  filter(Date >= last30 & Date < currentmonth)

#### Import tenth mile grid of the City of Dallas #####
grid_incidents <- st_read("data/dallas_tenthmilegrid.geojson") %>%
  st_transform(crs = 4269) %>%
  mutate(vgc_5year = lengths(st_intersects(., vgc5year)),
         vgc_12months = lengths(st_intersects(., vgc12months)),
         vgc_5yearAvg = vgc_5year/5,
         vgc_perch = (vgc_12months-vgc_5yearAvg)/vgc_5yearAvg)

#### Import  Dallas boundaries #####
DallasBoundary <- st_read("data/dallas_simpleboundary.geojson") %>%
  st_transform(crs = 4269)

dpd_beats <- st_read("data/dpd_beats.geojson") %>%
  st_transform(crs = 4269)

dpd_divisions <- st_read("data/dpd_divisions.geojson") %>%
  st_transform(crs = 4269)

incident_comp <- dpdIncidents %>%
    filter(crime %in% c("AGG ASSAULT - NFV", "MURDER & NONNEGLIGENT MANSLAUGHTER") & Firearm == 1) %>%
  filter((Year == 2022 & Month == month(last30)) |
           (Year == 2021 & Month == month(last30)) |
           (Year == 2021 & Month == month(12))) %>%
  st_drop_geometry(.) %>%
  group_by(Year, Month) %>%
  summarise(count = n())

incident_01 <- incident_comp %>%
  filter(Year == 2022 & Month == month(last30))

incident_02 <- incident_comp %>%
  filter(Year == 2021 & Month == month(12)) #month(last30)-month(1)

incident_03 <- incident_comp %>%
  filter(Year == 2021 & Month == month(last30))

incident_04 <- dpdIncidents %>%
    filter(crime %in% c("AGG ASSAULT - NFV", "ROBBERY-BUSINESS", "ROBBERY-INDIVIDUAL", "BURGLARY-RESIDENCE", "BURGLARY-BUSINESS", "MURDER & NONNEGLIGENT MANSLAUGHTER")) %>%
  filter((Year == 2022 & Month == month(last30))) %>%
  st_drop_geometry(.) %>%
  group_by(Year, Month) %>%
  summarise(count = n())
```

```{r, High Risk Cells, include = FALSE}
#### Import High Risk RTM Cells from most recent models
hr_directory <- paste0(datalib, "Safe Surroundings/04_Projects/RTM 2021/2021_Q3_RTM/Data/RTM_Q32021_Outputs.gpkg")

SE_HR <- st_read(hr_directory, layer = "SE_HighRisk") %>%
  st_transform(crs = 4269)
SC_HR <- st_read(hr_directory, layer = "SC_HighRisk") %>%
  st_transform(crs = 4269)
C_HR <- st_read(hr_directory, layer = "C_HighRisk") %>%
  st_transform(crs = 4269)
NC_HR <- st_read(hr_directory, layer = "NC_HighRisk") %>%
  st_transform(crs = 4269)
NE_HR <- st_read(hr_directory, layer = "NE_HighRisk") %>%
  st_transform(crs = 4269)
NW_HR <- st_read(hr_directory, layer = "NW_HighRisk") %>%
  st_transform(crs = 4269)
SW_HR <- st_read(hr_directory, layer = "SW_HighRisk") %>%
  st_transform(crs = 4269)

unique(dpdMini$division)

SC_Mark <- dpdMini %>%
  filter(division == "South Central") %>%
  .[SC_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

SE_Mark <- dpdMini %>%
  filter(division == "Southeast") %>%
  .[SE_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

NE_Mark <- dpdMini %>%
  filter(division == "Northeast") %>%
  .[NE_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

NW_Mark <- dpdMini %>%
  filter(division == "Northwest") %>%
  .[NW_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

C_Mark <- dpdMini %>%
  filter(division == "Central") %>%
  .[C_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

SW_Mark <- dpdMini %>%
  filter(division == "Southwest") %>%
  .[SW_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

NC_Mark <- dpdMini %>%
  filter(division == "North Central") %>%
  .[NC_HR, ] %>%
  mutate(HighRisk = "TRUE") %>%
  st_drop_geometry(.) %>%
  select(servnumid, HighRisk)

dpdMini <- bind_rows(SC_Mark, SE_Mark, NE_Mark, NW_Mark, C_Mark, SW_Mark, NC_Mark) %>%
  left_join(dpdMini, .)

rm(SC_Mark)
rm(SE_Mark)
rm(SW_Mark)
rm(C_Mark)
rm(NC_Mark)
rm(NE_Mark)
rm(NW_Mark)

month_01 <- as.character(month(last30, abbr = FALSE, label = TRUE))
month_02 <- as.character(month(last30 %m-% months(1), abbr = FALSE, label = TRUE))
year_01 <- as.character(year(last30))
year_02 <- as.character(year(last30-years(1)))

Division_VGC <- dpdMini %>%
  st_drop_geometry(.) %>%
  filter(vgc_flag == 1) %>%
  group_by(division) %>%
  summarise(count = n()) %>%
  mutate(perc = count/sum(count))

maxDivision <- filter(Division_VGC, perc == max(perc))

vgc_hr <- dpdMini %>%
  st_drop_geometry(.) %>%
  filter(vgc_flag == 1) %>%
  group_by(HighRisk) %>%
  summarise(count = n()) %>%
  filter(HighRisk == TRUE)
```


## `r paste(month_01, year_01)`

The violent crime rate for **`r paste(month_01, year_01)`** was **`r round(incident_01$count/(1304379/10000), digits = 3)`** per 10,000 residents.

In the month of **`r paste(month_01, year_01)`** there was a total of **`r incident_01$count`** violent gun crimes across the city. 

This is **`r ifelse(incident_01$count > incident_02$count, "an increase", "a decrease")` of `r paste0(round(((incident_01$count-incident_02$count)/incident_02$count)*100, digits = 1), "%")`** from **`r paste(month_02, year_01)`**, and **`r ifelse(incident_01$count > incident_03$count, "an increase", "a decrease")`** of **`r paste0(round(((incident_01$count-incident_03$count)/incident_03$count)*100, digits = 1), "%")`** from **`r paste(month_01, year_02)`**.

There were **`r nrow(filter(dpdMini, vgc_flag == 1 & tot_victims >= 4))`** incidents with 4 or more victims across the city.

**`r paste0(round(max(Division_VGC$perc)*100, digit = 1), "%")`** of all violent gun crimes in **`r paste(month_01, year_01)`** occurred in **`r maxDivision$division`** Division 

**`r paste0(round(vgc_hr$count/incident_01$count*100, digit = 1), "%", " (", vgc_hr$count, ")")`** of Violent Gun Crimes occurred within a cell identified as "High Risk" by our most recent Risk Terrain Modeling Analysis.

The reported U.S. violent crime rate includes murder, rape and sexual assault, robbery, and assault